---
title: "EDCS access and merging of TXT files"
author: "Petra Hermankova"
date: "27/11/2020"
output:
  html_document:
    theme: united
    toc: yes
    toc_float: true
    number_sections: true
    toc_depth: 2
    df_print: paged
---

```{r setup, include=FALSE, echo=FALSE}
require("knitr")
opts_knit$set(root.dir = "/home/petra/Github/ClaussSlaby_ETL/")

library(tidyverse)
library(data.table)
library(sdam)
library(jsonlite)
```

# Merging the CSV files into one dataframe
```{r}
# listing all tsv files
temp <- list.files(path= "../data/2020_12_allProvinces/", pattern = "*.tsv")
```

## Loading data from multiple TSV files

```{r}
setwd("~/Github/EDCS_ETL/data/2020_12_allProvinces")
# loading multiples files
EDCS <- temp %>% map_df(~read_tsv(., col_types = cols(.default = "c")))

setwd("/home/petra/Github/ClaussSlaby_ETL/")
```

```{r}
head(EDCS)
```

```{r}
names(EDCS)
```

# Cleaning attributes

Some of the attributes need to be streamlined as they contain either lists of values rathen than individual observations, but they have a form of a text string. They need to be converted to a list so they can be analysed computationally.

## Cleaning place attribute

Transforming string character into list for better manipulation.
```{r}
EDCS$place_list<- str_split(EDCS$place, " / ")
```

How many unique place names there is?
```{r}
length(unique(unlist(EDCS$place_list)))
```

## Cleaning 'dating from' attribute

1. Transforming string containing multiple values into a list.
2. Transforming character into numeric value.
```{r}
EDCS$`dating from`
```


## Cleaning 'dating to' attribute

1. Transforming string containing multiple values into a list.
2. Transforming character into numeric value.

## Cleaning 'status' attribute

```{r}
EDCS$status_list<- str_split(EDCS$status, "; ")
```

Ceategorisation based on the typology of keyword in the attribute status:

```{r}
inscription_type_list <- c("tituli sepulcrales", "tituli fabricationis", "inscriptiones christianae", "tituli sacri", "tituli possessionis", "tituli operum", "miliaria", "tituli honorarii", "carmina", "signacula", "diplomata militaria", "leges", "defixiones", "termini", "reges", "signacula medicorum", "senatus consulta")

inscribing_process_list <- c("sigilla impressa", "litterae erasae", "litterae in litura", "tesserae nummulariae")

status_notation_list<- c("viri", "tria nomina", "mulieres", "nomen singulare", "liberti/libertae", "milites", "Augusti/Augustae", "ordo senatorius", "servi/servae", "officium/professio", "ordo decurionum", "sacerdotes pagani", "praenomen et nomen", "ordo equester", "seviri Augustales", "sacerdotes christiani")
```

Division the status keywords into their own attributes, base on their purpose:
1. inscription type
2. notation of a societal status of a person on the inscription
3. details about process of inscribing, execution

```{r}
EDCS <- EDCS %>% 
  mutate(inscr_type = str_extract_all(pattern = paste(inscription_type_list, collapse="|"), string = EDCS$status)) %>% 
  mutate(status_notation = str_extract_all(pattern = paste(status_notation_list, collapse="|"), string = EDCS$status)) %>% 
  mutate(inscr_process = str_extract_all(pattern = paste(inscribing_process_list, collapse="|"), string = EDCS$status))
```

CONTINUE: need to convert character (0) to NA or eliminate.

```{r}
EDCS$status_notation

```

## Cleaning 'province' attribute

Some records contain multiple values provinces, separated by |, such as: 

`Belgica | Germania inferior`
`Pannonia inferior | Pannonia superior`
`Moesia inferior | Moesia superior`

```{r}
EDCS$province_list <- str_split(EDCS$province, " \\| ")
```

Listing all the provinces and their variants:
```{r}
unique(EDCS$province_list)
```




















# Saving to Scienc Data



```{r, echo=FALSE}
mycred_secret<- readLines("~/mysecret.txt")
```

```{r}
EDCS_json <- jsonlite::toJSON(EDCS, auto_unbox=TRUE)
write(EDCS_json, file="EDCS_merged_2020-12-09.json")
request("EDCS_merged_2020-12-09.json", path="/sharingout/648597@au.dk/SDAM_root/SDAM_data/EDCS/public",
        method="PUT", cred=c(mycred_secret[1], mycred_secret[2]))
```

```{r, echo=FALSE}
remove(mycred_secret)
```
