---
title: "EDCS cleaning of the epigraphic text"
author: "Petra Hermankova"
date: "10/12/2020"
output:
  html_document:
    theme: united
    toc: yes
    toc_float: true
    number_sections: true
    toc_depth: 2
    df_print: paged
---

```{r setup, include=FALSE, echo=FALSE}
require("knitr")
#opts_knit$set(root.dir = "/home/petra/Github/EDCS_ETL/")

library(tidyverse)
library(jsonlite)
library(leaflet)
```

## Loading data
Load the dataset, if you have Sciencedata.dk credentials

Make a list and tibble from the downloaded dataset
```{r}
EDCS <- jsonlite::fromJSON("https://sciencedata.dk/public/1f5f56d09903fe259c0906add8b3a55e/EDCS_merged_cleaned_attrs_2021-03-01.json")
```

# Cleaning epigraphic text for tidy text mining analysis (word and sentence centered)

*Aim:* The main purpose of this script is to clean large collections of epigraphic texts all at once in order to create cleaned texts ready for text mining analysis. The output clean texts can be used for a) word centered text mining, also known as the tidytext approach (https://www.tidytextmining.com/) or for b) sentence centered text mining, as part of the Natural Language Processing (https://en.wikipedia.org/wiki/Natural_language_processing).

The presented cleaning process is designed as generic, fairly modular and fully customisable. Therefore, it can be with some modification used to any epigraphic corpus. Ample examples are provided to illustrate individual parts of the process, so anyone familiar with _Regular Expressions_ and _basic understanding of R_ can build their own cleaning function or modify the existing ones.

The final output of the cleaning function depends on which of the individual cleaning blocks will be used and in what sequence they will run. Each individual cleaning block represents one pattern occurring repeatedly in the text that can be searched for and modified, depending on the intended outcome. All the cleaning steps are dependent on the characteristics of the original dataset, therefore familiarity with the original dataset prior the cleaning process is recommended. Each dataset can have a different set of symbols and characters to be cleaned, thus, the cleaning blocks should be adjusted accordingly.

I have created three categories of cleaning blocks, closely linked with the methodological approach and the purpose of the cleaning process:

1. `Conservative cleaning model` producing a text as close to the original as possible
2. `Interpretive cleaning model` producing a text enriched with interpretations of the corpus editor
3. Generic cleaning of patterns common for both previous categories

_Structure of a cleaning block:_

Each of the cleaning blocks maintains the same structure, using Regular expressions to find and replace the searched term or pattern.

```regexpatternname <- c("regexpattern", "substitutionpattern")```
## 1. Cleaning blocks for the conservative model
*The aim of this model is to produce a clean text that is as close to the original text of an inscription as possible, without any editorial input.*
The cleaned output of the conservative model will be as close to the original text of an inscription as possible. In most cases it should resemble a _diplomatic edition_ of epigraphic text with spaces between words, lowercase letters, eliminated brackets and non-utf compliant symbols. The interpretive restoration, substitutions or any changes of the text as appear in the dataset, done by the editor of the epigraphic corpus, are eliminated from the conservative model.
### 1.1. Expanded abbreviations
**Aim:** All expanded abbreaviations that are in the parenthesis () will be eliminated from the clean text (substituted with "").
* Example before cleaning: ```Αὐρ(ήλιος) Οὐαλέριος```
* Example after cleaning: ```Αὐρ Οὐαλέριος```
```{r}
expanded_abbreviations_conservative <- c("\\([^(]*\\)", "")
```

### 1.2. Suppresion of a text with superscripts

**Aim:** All supressions that are in the curly braces {} followed by one or more superscript digits will be eliminated from the clean text (substituted with "").

**!!!** It is crucial that block `suppresion_conservative` does not precede block `suppresion_superscripts_conservative`, otherwise the Regex pattern would not clean the text properly. This particular pattern is common for the PHI dataset and may  or may not appear in other datasets.

* Example before cleaning: ```ἱερεὺς ληφθὶς ὑπὰ {²⁶ὑπὸ}²⁶ τῶν βαρβάρων ```
* Example after cleaning: ```ἱερεὺς ληφθὶς ὑπὰ  τῶν βαρβάρων ```

```{r}
suppresion_superscripts_conservative <- c("{[^}]*}[⁰¹²³⁴⁵⁶⁷⁸⁹]+", "")
```

### 1.3. Suppresion of a text

**Aim:** All curly braces {} will be eliminated from the clean text (substituted with ""), while the contents of the braces will remain in the text.

**!!!** It is crutial that block `suppresion_conservative` does not precede block `suppresion_superscripts_conservative`, otherwise the Regex pattern would not clean the text properly.

* Example before cleaning: ```Σεβαστοῦ υἱοῦ {θ̣εοῦ Σεβαστοῦ} τύχης  ```
* Example after cleaning: ```Σεβαστοῦ υἱοῦ θ̣εοῦ Σεβαστοῦ τύχης  ```

```{r}
suppresion_conservative <- c("[\\{*\\}]", "")
```

### 1.4. Restoration

**Aim:** All restoration that are in the square brackets [] will be eliminated from the clean text (substituted with "").

**!!!** Beware that by eliminating the contents of the brackets you may loose some context - use at your own discretion.

* Example before cleaning: ```[Ν]ανα Ἕλληνο̣[ς] θυγάτηρ καὶ ἡ ἑτέρα [γυνὴ]```
* Example after cleaning: ```ανα Ἕλληνο θυγάτηρ καὶ ἡ ἑτέρα```

```{r}
restoration_conservative <- c("\\[[^[]*\\]", "")
```

### 1.5. Substitution

**Aim:** All substitutions that are in the angular brackets <> will be eliminated from the clean text (substituted with "").

**!!!** Beware that by eliminating the contents of the brackets you may loose some context - use at your own discretion.

* Example before cleaning: ```κωρο<ν Ἀ>ντιόχ<ου> ἡ πατρὶς τειμῆ<ς>```
* Example after cleaning: ```κωρο ντιόχ ἡ πατρὶς τειμῆς```

```{r}
substitution_conservative <- c("\\<[^<]*\\>", "")
```

### 1.6. Substitution in EDH dataset

**Aim:** All  sustitutions following the pattern "A=B" will be cleaned thw following way: B remain in the text and the equal sign and A will be eliminated from the clean text.

**!!!** Beware that by eliminating the brackets you may loose some information about the preservation of the text - use at your own discretion. The `substitution_edh_interpretive` should be run before `substitution_interpretive` block, otherwise the Regex pattern would not clean the text properly. The `substitution_interpretive` block will clean the angular brackets in the next step.

* Example before cleaning: ```pos<u=I>erunt bene merenti```
* Example after cleaning: ```pos<I>erunt bene merenti```

```{r}
substitution_edh_conservative <- c("([α-ωΑ-Ωa-zA-Z])=([α-ωΑ-Ωa-zA-Z])", "\\2")
```

## 2. Cleaning blocks for the interpretive model

*The aim of this model is to produce a clean text that is enriched with interpretations of the original text as published by the editor of the corpus. The editorial interpretations include abbreviations, restorations, substitutions and suppresions of the text.*

The output of the interpretive model will produce an epigraphic text with as many editorial suggestions, restorations, corrections, and improvements as possible to provide as much possible contents of the inscription as possible. The brackets and non-utf compliant symbols will be eliminated from the `interpretive model`.

### 2.1. Expanded abbreviations

**Aim:** All parenthesis () will be eliminated from the clean text (substituted with ""), while the contents of the parenthesis will remain in the text.

* Example before cleaning: ```Αὐρ(ήλιος) Οὐαλέριος```
* Example after cleaning: ```Αὐρήλιος Οὐαλέριος```

```{r}
expanded_abbreviations_interpretive <- c("[\\(*\\)]", "")
```

### 2.2. Suppresion of a text with superscripts

**Aim:** Contents found within curly braces {} followed by one or more superscript digits will substitute the word immediately preceding the curly braces with the word contained in the curly braces and the braces will be eliminated, see example. Note: The cleaning block will not work if there is no text preceeding the curly braces (the pattern will be skipped).

**!!!** This particular pattern is common for the PHI dataset and may  or may not appear in other datasets. It is recommended to run the ```suppresion_keep_interpretive``` or ```suppresion_remove_interpretive``` block after ```suppresion_superscripts_interpretive``` block, otherwise the Regex pattern would not clean the text properly.

* Example before cleaning: ```ἱερεὺς ληφθὶς ὑπὰ {²⁶ὑπὸ}²⁶ τῶν βαρβάρων ```
* Example after cleaning: ```ἱερεὺς ληφθὶς ὑπὸ τῶν βαρβάρων ```

```{r}
suppresion_superscripts_interpretive <- c(" [^ ]+ \\{([⁰¹²³⁴⁵⁶⁷⁸⁹]+)([^}]+)\\}\\1", " \\2")
```

### 2.3. Suppresion of a text

**Aim:** All curly braces {} will be eliminated from the clean text (substituted with ""), while the contents of the braces will remain in the text.

**!!!** It is crutial that block ```suppresion_keep_interpretive``` or ```suppresion_remove_interpretive``` does not precede block `suppresion_superscripts_interpretive`, otherwise the Regex pattern would not clean the text properly. Due to ambiguous use of {} by editors of epigraphic corpora, the exact usage depends on the specific dataset and the way the curly braces were used. Therefore, two options how to handle curly braces are provided: If you wish to keep the text within the curly braces and remove the braces, use ```suppresion_keep_interpretive``` block. If you wish to remove the text in the braces and the braces, use ```suppresion_remove_interpretive``` block.

* Example before cleaning: ```θ̣εοῦ Σεβαστοῦ υἱοῦ {θ̣εοῦ Σεβαστοῦ} τύχης  ```
* Example after cleaning (keep text): ```θ̣εοῦ Σεβαστοῦ υἱοῦ θ̣εοῦ Σεβαστοῦ τύχης```
* Example after cleaning (remove text): ```θ̣εοῦ Σεβαστοῦ υἱοῦ  τύχης```

```{r}
suppresion_keep_interpretive <- c("[\\{*\\}]", "")
```

OR if you wish to remove the contents of the braces

```{r}
suppresion_remove_interpretive <- c("{[^}]*}", "")
```

### 2.4. Restoration

**Aim:** All square brackets [] will be eliminated from the clean text (substituted with ""), while the contents of the brackets will remain in the text.

**!!!** Beware that by eliminating the brackets you may loose some information about the preservation of the text - use at your own discretion.

* Example before cleaning: ```[Ν]ανα Ἕλληνο̣[ς] θυγάτηρ καὶ ἡ ἑτέρα [γυνὴ]```
* Example after cleaning: ```Νανα Ἕλληνο̣ς θυγάτηρ καὶ ἡ ἑτέρα γυνὴ```

```{r}
restoration_interpretive <- c("[\\[*\\]]", "")
```

### 2.5. Substitution

**Aim:** All angular brackets <> will be eliminated from the clean text (substituted with ""), while the contents of the brackets will remain in the text.

**!!!** Beware that by eliminating the brackets you may loose some information about the preservation of the text - use at your own discretion.  

* Example before cleaning: ```κωρο<ν Ἀ>ντιόχ<ου> ἡ πατρὶς τειμῆ<ς>```
* Example after cleaning: ```κωρον Ἀντιόχου ἡ πατρὶς τειμῆς```

```{r}
substitution_interpretive <- c("[\\<*\\>]", "")
```

### 2.6. Substitution in the EDH dataset

**Aim:** All  sustitutions following the pattern "A=B" will be cleaned the following way: "A" will remain in the text and the equal sign and "B" will be eliminated from the clean text.

**!!!** The `substitution_edh_interpretive` should be run before `substitution_interpretive` block, otherwise the Regex pattern would not clean the text properly. The `substitution_interpretive` block will clean the angular brackets in the next step.

* Example before cleaning: ```pos<u=I>erunt bene merenti```
* Example after cleaning: ```pos<u>erunt bene merenti```

```{r}
substitution_edh_interpretive <- c("([α-ωΑ-Ωa-zA-Z])=([α-ωΑ-Ωa-zA-Z])", "\\1")
```

### 2.7. Substitution in the EDH dataset with missing first part - interpretive

**Aim:** All  substitutions following the pattern "<=B>" will be cleaned the following way: the equal sign will be eliminated from the clean text and the and "B" and brackets will remain for further cleaning.

**!!!** The `substitution_edh_interpretive_missing` should be run before `substitution_interpretive` block, otherwise the Regex pattern would not clean the text properly. The `substitution_interpretive` block will clean the angular brackets in the next step.

* Example before cleaning: ```Dion<=Y>si ```
* Example after cleaning: ```Dion<Y>si ```

```{r}
substitution_edh_interpretive_missing <- c("([α-ωΑ-Ωa-zA-Z])*=([α-ωΑ-Ωa-zA-Z])", "\\2")
```
### 2.8. Substitution in the EDH dataset with missing first part - conservative

**Aim:** All  substitutions following the pattern "<=B>" will be cleaned the following way: the equal sign and "B" will be eliminated from the clean text  and the the brackets will remain for further cleaning.

**!!!** The `substitution_edh_interpretive_missing` should be run before `substitution_interpretive` block, otherwise the Regex pattern would not clean the text properly. The `substitution_interpretive` block will clean the angular brackets in the next step.

* Example before cleaning: ```Dion<=Y>si ```
* Example after cleaning: ```Dion<>si ```

```{r}
substitution_edh_conservative_missing <- c("([α-ωΑ-Ωa-zA-Z])*=([α-ωΑ-Ωa-zA-Z])", "\\1")
```


## 3. The generic text cleaning

*The aim of the generic cleaning is to strip the epigraphic text any non-utf compliant symbols and characters that do not adhere to the principles of a quantitative text mining.*

The cleaning blocks in this section represent common patterns appearing in any epigraphic text, such as interpunction, lacunas or other representations of an empty space, various editorial notes and comments in the text itself, that are not relevant to the text mining, erasures, numerals, and several specific unicode symbols appearing in the original text. Depending on the characteristics of the originahal dataset and the intended outcome, anyone can change individial cleaning blocks to better fit their needs. Through testing is, however, strongly recommended!

### 3.1. Lacuna 1

**Aim:** All square brackets [] containing one or more "— " will be eliminated from the clean text (substituted with "").

**!!!** The block ```lacuna1``` should be run before ```restoration_conservative``` and ```restoration_interpretive``` blocks, otherwise the Regex pattern would not clean the text properly. Note: If there is a text within the square bracket, e.g. ```προύχον[τος — — —]``` the block ```lacuna1``` will skip the pattern. However, the block ```restoration_interpretive``` will eliminate the square brackets, the script ```interpunction_symbols``` will clean the "—" and the script ```multi_whitespace``` will eliminate the extra whitespaces. Therefore the blocks should be used in combination and in the indicated sequence: (1)```restoration_interpretive```, (2)```interpunction_symbols``` and (3)```multi_whitespace```.

* Example before cleaning: ```[— — —]ης θεῷ Φοίβῳ```
* Example after cleaning: ```ης θεῷ Φοίβῳ```

```{r}
lacuna1 <- c("\\[[— ]+\\]", "")
```

### 3.2. Lacuna 2

**Aim:** All square brackets [] containing one or more "." will be eliminated from the clean text (substituted with "").

**!!!** The block ```lacuna2``` should be run before ```restoration_conservative``` and ```restoration_interpretive``` blocks, otherwise the Regex pattern would not clean the text properly. Note: If there is a text within the square bracket, e.g. ```προύχον[τος...]``` the block ```lacuna2``` will skip the pattern. However, the block ```restoration_interpretive``` will eliminate the square brackets, the script ```interpunction_symbols``` will clean the "." and the script ```multi_whitespace``` will eliminate the extra whitespaces. Therefore the blocks should be used in combination and in the indicated sequence: (1)```restoration_interpretive```, (2)```interpunction_symbols``` and (3)```multi_whitespace```.

* Example before cleaning: ```[․․]ω Διὶ καὶ Ἥρᾳ```
* Example after cleaning: ```ω Διὶ καὶ Ἥρᾳ```

```{r}
lacuna2 <- c("\\[[․]+\\]", "")
```

### 3.3. Vacat

**Aim:** All instances of the following strings "vacat, vac, vac., v." will be replaced by a space (substituted with " "). If there is any extra whitespace, it will be cleaned by ```multi_whitespace``` block in the following steps.

**!!!** If your datasets contains latin inscriptions, you may want to check whether the ```vacat``` block is not eliminitating more words than anticipated, e.g. words containing string "vacat" or "vac". If so, adjust the cleaning block accordingly, i.e. remove "vac", or don't use it.

* Example before cleaning: ```Ἡρακλείδα vacat χαῖρε.```
* Example after cleaning: ```Ἡρακλείδα    χαῖρε.```

```{r}
vacat <- c("(vacat|vac|vac\\.|v\\.)", " ")
```

### 3.4. Editorial notes

**Aim:** All instances of the editorial notes in parenthesis such as (vel sim.) will be replaced by a space (substituted with " "). If there is any extra whitespace, it will be cleaned by ```multi_whitespace``` block in the following steps.

**!!!** The ```editorial_notes``` block should run before the ```expanded_abbreviations_conservative``` and ```expanded_abbreviations_interpretive``` blocks, otherwise the Regex pattern would not clean the text properly.

* Example before cleaning: ```Ἥρωι (vel sim.) Καλλισθένης```
* Example after cleaning: ``Ἥρωι   Καλλισθένης```

```{r}
editorial_notes <-c("\\(vel sim.\\)", " ")
```

### 3.5. New line

**Aim:** All instances of in-line symbol for new line (|) will be eliminated (substituted with "").

* Example before cleaning: ```Λάμπρη// Τ̣ελεσήνορ|ος γυνή```
* Example after cleaning: ```Λάμπρη Τ̣ελεσήνορος γυνή```

```{r}
new_line <- c("[\\||\\/|\\/\\/]", "")
```

### 3.6. Split word over two lines

**Aim:** All instances of words split between two lines with a dash (-) will be eliminated (substituted with "").

* Example before cleaning: ```ἀρχιερέως καὶ εὐποσιάρ-\nχου μηνὸς```
* Example after cleaning: ```ἀρχιερέως καὶ εὐποσιάρχου μηνὸς```

```{r}
split_word_multiline <- c("-\\n", "")
```

### 3.7. Erasure empty

**Aim:** All instances of erased text (〚—〛) will be replaced by a space (substituted with " "). If there is any extra whitespace, it will be cleaned by ```multi_whitespace``` block in the following steps.

* Example before cleaning: ```Ἀρτέμιδι 〚— — —〛 ἐπηκόοις.```
* Example after cleaning: ```Ἀρτέμιδι  ἐπηκόοις.```

```{r}
erasure_empty <- c("〚[— ]+〛", " ")
```

### 3.8. Erasure with new text

**Aim:** All instances of double brackets for erasures (〚 〛) will be eliminated (substituted with "") and the contents of the double brackets will be preserved as part of the clean text.

* Example before cleaning: ```Ἀμύντωρ Νουμηνίου 〚χαῖρε〛. καὶ ἡ γυνὴ αὐτοῦ```
* Example after cleaning: ```Ἀμύντωρ Νουμηνίου χαῖρε. καὶ ἡ γυνὴ αὐτοῦ```

```{r}
erasure_new_text <- c("[〚〛]", "")
```

### 3.9. Dubious dot subscript

**Aim:** All instances of the dubious reading marked by the subscrit dot (unicode 0323) will be eliminated (substituted with "").

**!!!** The ```dubious_dot_subscript``` block should happen as first step of the cleaning, otherwise the letters might shift and the Regex pattern would not clean the text properly.

* Example before cleaning: ``` Ἀ̣πό̣λ̣λ̣ωνος```
* Example after cleaning: ``` Ἀπόλλωνος```

```{r}
dubious_dot_subscript <- c("\u{0323}", "")
```

### 3.10. Interpunction symbols

**Aim:** All instances of listed interpunction symbols (,.!-—#%^&\*/~:;) will be replaced by a space (substituted with " "). If there is any extra whitespace, it will be cleaned by ```multi_whitespace``` block in the following steps.

**!!!** If you wish to keep sentence separators, such as dots at the bottom of the line, use ```interpunction_keep_sentences``` or elimininate the sentence separators you want to keep in your text from the cleaning block ```interpunction_keep_sentences```.

* Example before cleaning: ```Φιλήτη # θεᾷ Μαλοφόρῳ``` or ```κεῖμαι πρόμοιρος Ἑρμογένης τυμβευθείς. /ἀγὼν```
* Example after cleaning: ```Φιλήτη  θεᾷ Μαλοφόρῳ``` or ```κεῖμαι πρόμοιρος Ἑρμογένης τυμβευθείς   ἀγὼν```
* Example after cleaning (keep sentences): ```κεῖμαι πρόμοιρος Ἑρμογένης τυμβευθείς.  ἀγὼν```

```{r}
interpunction_symbols <- c("[=\\+\\,|\\.|․|:|⋮|⁙|;|!|\\-|—|–|#|%|\\^|&|\\~|@]", " ")
```
OR

if you wish to preserve sentence separators, such as dots

```{r}
interpunction_keep_sentences <- c("[!|\\-|—|–|#|%|\\^|&|\\*|~|@\\+]", " ")
```

### 3.11. Superscript numbers

**Aim:** All instances of superscripted numbers will be eliminated (substituted with "").

**!!!** The ```superscript_numbers``` should not be run before the ```suppresion_superscripts_conservative``` or ```suppresion_superscripts_interpretive``` block, otherwise the Regex pattern would not clean the text properly.

* Example before cleaning: ```Αὐρ(ήλιος) Διονύσιος #⁵⁶ βʹ #⁵⁶```
* Example after cleaning: ```Αὐρ(ήλιος) Διονύσιος # βʹ #```

```{r}
superscript_numbers <- c("[⁰¹²³⁴⁵⁶⁷⁸⁹]+", "")
```

### 3.12. Epigraphic symbols

**Aim:** All instances of the listed specialised epigraphic symbols, such as the haedera (❦), will be eliminated (substituted with "").

* Example before cleaning: ```ἀγαθῆι ❦ τύχηι```
* Example after cleaning: ```ἀγαθῆι   τύχηι```

```{r}
epigraphic_symbols <-c ("[❦|·|∙|𐆖|⏑|⏓|⏕]", "")
```

### 3.13. Uncertainty symbols

**Aim:** All instances of th elisted symbols marking uncertainty (?) will be eliminated (substituted with "").

* Example before cleaning: ```χαῖρε?```
* Example after cleaning: ```χαῖρε```

```{r}
uncertainty_symbols <-c ("[\\?]", "")
```

### 3.14. End of line

**Aim:** All instances of end of line symbol (\n) will be replaced by space (substituted with " ").

* Example before cleaning: ```καὶ ἄρξαντα\nτοῦ κοινοῦ```
* Example after cleaning: ```καὶ ἄρξαντα τοῦ κοινοῦ```

```{r}
end_line <- c("\\n", " ")
```

### 3.15. Extra blank space

**Aim:** All instances of extra blank space (" ") will be replaced by space (substituted with " ").

* Example before cleaning: ```ἀγαθῆι   τύχηι.```
* Example after cleaning: ```ἀγαθῆι   τύχηι.```

```{r}
extra_blank <- c("[ ]+", " ")
```

### 3.16. Multi-whitespace

**Aim:** All instances of more then one whitespace "  " next to each other will be eliminated (substituted with "").

**!!!** The ```multi_whitespace``` should run as the second last cleaning block to ensure all redundant white spaces are cleaned from the text.

* Example before cleaning: ```Ἡρακλείδα    χαῖρε.```
* Example after cleaning: ```Ἡρακλείδα χαῖρε.```

```{r}
multi_whitespace <- c("\\s+", " ")
```

### 3.17. Trailing and leading whitespace

**Aim:** All instances of whitespace " " at the beginning and end of the line will be eliminated (substituted with "").

**!!!** The ```whitespace_endline``` should run as the last cleaning block to ensure all redundant white spaces are cleaned from the text.

* Example before cleaning: ``` χαῖρε ```
* Example after cleaning: ```χαῖρε```

```{r}
whitespace_endline <- c("(^\\s|\\s$)", "")
```

### 3.18. Editorial comments in Latin alphabet

**Aim:** All instances of editorial comments in Latin alphabet that are enclosed in curly braces {} with superscript numbers will be eliminated (substituted with "").

**!!!** If your dataset contains Latin inscriptions, use this block with caution. Verify first, that running the block  does not eliminate any necessary information or text. This block has been specifically designed for the interpretive cleaning of the PHI Greek Inscription dataset and it should run before ```suppresion_superscripts_interpretive``` and ```suppresion_interpretive``` blocks, otherwise the Regex pattern would not clean the text properly.

* Example before cleaning: ```ἀγαθῆι τύχηι. {²in parte inferiore altera manu incisa est:}² ὑπὲρ τῆς τοῦ```
* Example after cleaning: ```ἀγαθῆι τύχηι. ὑπὲρ τῆς τοῦ```

```{r}
editorial_comments_latin <- c("\\{([⁰¹²³⁴⁵⁶⁷⁸⁹]+)([a-zA-Z0-9][^}]+)\\}\\1", "")
```

### 3.19. Arabic numerals

**Aim:** All instances of arabic numerals (0-9) will be eliminated (substituted with "").

**!!!** If your dataset contains arabic numerals that you would like to keep, use this block with caution. Verify first, that running the block does not eliminate any necessary information or text. This block has been specifically designed for the interpretive cleaning of the PHI Greek Inscription dataset and it should run before ```multi_whitespace``` and ```whitespace_endline``` blocks, otherwise the Regex pattern would not clean the text properly.

* Example before cleaning: ```ἡ γυνὴ αὐτοῦ ΦιλΙ̣ 4 5 καὶ```
* Example after cleaning: ```ἡ γυνὴ αὐτοῦ ΦιλΙ καὶ```

```{r}
arabic_numerals <- c("[0-9]+", "")
```

### 3.20 Unclosed brackets

**Aim:** All instances of unclosed brackets will be eliminated (substituted with "").

**!!!** Use the `unclosed_brackets` block immediately before ```multi_whitespace``` and ```whitespace_endline``` blocks, otherwise the Regex pattern would not clean the text properly.

* Example before cleaning: ```ummio isenna Xv [```
* Example after cleaning: ```ummio isenna Xv ```

```{r}
unclosed_brackets <- c("[\\[|\\{|\\(|\\)|\\}|\\]]", "")
```


## EDCS specific substitutions

### 4.1 Blank within line represented by number 3 in square brackets []

**Aim:** All instances of number three in square brackets will substituted with "[-]".

* Example before cleaning: ```d[3]ad```
* Example after cleaning: ```d[-]ad```

```{r}
edcs_number_three_both <- c("\\[3\\]", "[-]")
```

### 4.2 Blank within line represented by number 3 in square brackets closed from the right side

**Aim:** All instances of number three in right square brackets will substituted with "-]".

* Example before cleaning: ```[mina3]ad```
* Example after cleaning: ```[mina-]ad```

```{r}
edcs_number_three_right <- c("3\\]", "-]")
```

### 4.3 Blank within line represented by number 3 in square brackets closed from the left side

**Aim:** All instances of number three in left square brackets will substituted with "[-".

* Example before cleaning: ```ad locuti[3nis]```
* Example after cleaning: ```ad locuti[-nis]```

```{r}
edcs_number_three_left <- c("\\[3", "[-")
```


### 4.4 Blank line represented by number 6 in square brackets []

**Aim:** All instances of number six in left square brackets will substituted with "[-".

* Example before cleaning: ```ad locuti [6] ergo```
* Example after cleaning: ```ad locuti [-] ergo```

```{r}
edcs_number_six_both <- c("\\[6\\]", "[-]")
```


### 4.5 Blank space represented by number 1

**Aim:** All instances of number one will substituted with " ".

* Example before cleaning: ```nobbiliss1mis Caessaribus```
* Example after cleaning: ```nobbiliss mis Caessaribus```

```{r}
edcs_number_one <- c("[1]", " ")
```

### 4.6 Quotes

**Aim:** All instances of quotes one will substituted with " ".

* Example before cleaning: ```"ave" Caessaribus```
* Example after cleaning: ``` ave  Caessaribus```

```{r}
edcs_quotes <- c("\u{0022}", " ")
```

### 4.7 Backslashes

**Aim:** All instances of backslashes one will substituted with "".

* Example before cleaning: ```Eros Cai Epidi \\ sum```
* Example after cleaning: ```Eros Cai Epidi  sum```



```{r}
edcs_backslashes <- c("\u{005C}\u{005C}", " ")
```


### 4.8 Latin Enclitics

**Aim:** All instances of Latin eclitics -que will be separated from the word by space.

* Example before cleaning: ```libertatisque```
* Example after cleaning: ```libertatis que```

```{r}
edcs_que <- c("(\\w+)(que)\\b", "\\1 \\2")
```

### 4.9 Roman numerals + vir

**Aim:** All instances of Roman numerals preceeding word 'vir' will be separated from the word by space.

* Example before cleaning: ```IIIIviribus```
* Example after cleaning: ```IIII viribus```

```{r}
edcs_vir <- c("([I|V|X])(vir*)", "\\1 \\2")
```



## EDCS specific cleaning

### Removing redundant onformation from `inscription`

1) Attribute `inscription` contains over 1000 of notes to dating of the inscription that needs to be matched with Regexes, extracted to separate attribute `dating_notes` and then remove from the text of an inscription and saved as a new attribute `inscription_stripped`.

*Example of strings to be removed*

`to 276;   b:  276 to 282" \n\n \n \n`
`to 367;   b:  ;   c:  367 to 375;   d:  387 to 388" \n\n \n \n`

2) Attribute `inscription` contains references to DOI that are extracted to separate attribute `other_notes` and removed from the text of an inscription and saved as a new attribute `inscription_stripped2`

*Example of strings to be removed*

` \n\n10.3406/mefr.1975.1032`

3) Attribute `inscription` contains modern notes that are not part of the ancient inscription and need to be cleaned out.

*Example of strings to be removed*

`According the proposed supplements of the inscription the date is 162 or 163 like EDCS-28500161 from the year 162/3 or EDCS-29100030 from the year 162.`



```{r, stripping the text of an inscription}
EDCS_clean <- EDCS %>%
   mutate(dating_notes = str_extract(EDCS$inscription, "^[to|b:].*\\n\\n")) 
EDCS_clean <- EDCS_clean %>%
   mutate(inscription_stripped = str_replace(EDCS$inscription, "^[to|b:].*\\n\\n([^.*])", "\\1")) 
EDCS_clean <- EDCS_clean %>%
   mutate(dating_notes2 = str_extract(EDCS_clean$inscription_stripped, "^to\\s{1,2}\\d{1,4}.*")) 
EDCS_clean <- EDCS_clean %>%
   mutate(inscription_stripped2 = str_replace(EDCS_clean$inscription_stripped, "^to\\s{1,2}\\d{1,4}.*", "")) 
EDCS_clean <- EDCS_clean %>%
   mutate(other_notes = str_extract(EDCS_clean$inscription_stripped2, "\\n.*10\\.\\d{1,6}.+$")) 
EDCS_clean <- EDCS_clean %>%
   mutate(inscription_stripped3 = str_replace(EDCS_clean$inscription_stripped2, "\\n.*10\\.\\d{1,6}.+$","")) 
EDCS_clean <- EDCS_clean %>%
   mutate(other_notes2 = str_extract(EDCS_clean$inscription_stripped3, "\\s*http.*"))
EDCS_clean <- EDCS_clean %>%
   mutate(inscription_stripped4 = str_replace(EDCS_clean$inscription_stripped3, "\\s*http.*","")) 
EDCS_clean <- EDCS_clean %>%
   mutate(other_notes3 = str_extract(EDCS_clean$inscription_stripped4, "\\ncomment\\n.*")) %>%
   mutate(inscription_stripped5 = str_replace(EDCS_clean$inscription_stripped4, "\\ncomment\\n.*","")) 
EDCS_clean <- EDCS_clean %>%
   mutate(other_notes4 = str_extract(EDCS_clean$inscription_stripped5, "\\n\\n(?!lauce|la s\\(olvit\\)).*")) %>%
   mutate(inscription_stripped6 = str_replace(EDCS_clean$inscription_stripped5, "\\n\\n(?!lauce|la s\\(olvit\\)).*","")) 
EDCS_clean <- EDCS_clean %>%
   mutate(other_notes5 = str_extract(EDCS_clean$inscription_stripped6, "\\[Manfred.*|\\[Werner.*\\]")) %>%
   mutate(inscription_stripped7 = str_replace(EDCS_clean$inscription_stripped6, "\\[Manfred.*|\\[Werner.*\\]","")) 

EDCS_clean <- EDCS_clean  %>% 
   mutate(inscription_stripped8 = str_replace(EDCS_clean$inscription_stripped7, "^ \\n \\n","")) # eliminating empty lines with spaces at the beginning
EDCS_clean <- EDCS_clean %>%
   mutate(inscription_stripped9 = str_replace(EDCS_clean$inscription_stripped8, "\\n$","")) # eliminating empty lines at the end
EDCS_clean <- EDCS_clean %>%
   mutate(inscription_stripped10 = str_replace(EDCS_clean$inscription_stripped9, "^ \\n\\n","")) # eliminating empty lines at the beginning

EDCS_clean <- EDCS_clean %>%
   mutate(other_notes6 = str_extract(EDCS_clean$inscription_stripped10, "\\n.*")) %>%
   mutate(inscription_stripped11 = str_replace_all(EDCS_clean$inscription_stripped10, "\\n.*","")) 
EDCS_clean <- EDCS_clean %>%
   mutate(other_notes7 = str_extract(EDCS_clean$inscription_stripped11, "\\bThe .*")) 
EDCS_clean <- EDCS_clean %>%
   mutate(inscription_stripped12 = str_replace(EDCS_clean$inscription_stripped11, "\\bThe .*","")) 
EDCS_clean <- EDCS_clean %>%
   mutate(other_notes8 = str_extract(EDCS_clean$inscription_stripped12, ".* the .*")) 

EDCS_clean <- EDCS_clean %>%
   mutate(inscription_stripped13 = str_replace(EDCS_clean$inscription_stripped12, ".* the .*","")) 

EDCS_clean <- EDCS_clean %>%
   mutate(inscription_stripped_final = str_replace(EDCS_clean$inscription_stripped13, "\\r+",""))

```

#### How many inscriptions were pre-cleaned: 

```{r}
Text_with_dating <- EDCS_clean %>% 
  dplyr::filter(notes_dating != " " )
Text_with_dating

Text_with_comments <- EDCS_clean %>% 
  filter(notes_comments != "     ")
Text_with_comments

Text_with_references <- EDCS_clean %>% 
  filter(notes_references != " " )
Text_with_references
```

```{r}
edcs_debug_date <- Text_with_dating %>% 
  select(`EDCS-ID`, inscription, Comment, `dating from`, `dating to`)

edcs_debug_comments <- Text_with_comments %>% 
  select(`EDCS-ID`, inscription, Comment)

edcs_debug_biblio <- Text_with_references %>% 
  select(`EDCS-ID`, inscription, Comment)

write_csv(edcs_debug_date, "../output/EDCS_debug_date.csv")
write_csv(edcs_debug_comments, "../output/EDCS_debug_comments.csv")
write_csv(edcs_debug_biblio, "../output/EDCS_debug_biblio.csv")
```



```{r}
precleaning <-  
  nrow(Text_with_dating) +
  nrow(Text_with_comments) +
  nrow(Text_with_references)

precleaning
precleaning/(nrow(EDCS)/100)
```


# Building cleaning functions for specific datasets

When we have established the individual building blocks, we can put them together in the right sequence and build a cleaning function in R for conservative and interpretive models.

### Conservative model for text mining analysis
*Aim:* to have a clean text enriched without any editorial interpretations and reconstructions of the text (to have a clean text that is as close to the original inscription as preserved on the medium). 

The output of the function will consist of words separated by one space, so the data is ready for tidytext analysis. No interpunction will be left in the text.

```{r}
cleaning_conservative_edcs <- function(epigraphic_dataset){
  clean_text <- gsub(pattern=dubious_dot_subscript[1], replacement=dubious_dot_subscript[2], x=epigraphic_dataset, perl=TRUE)
  clean_text <- gsub(pattern=edcs_number_three_both[1], replacement=edcs_number_three_both[2], x=clean_text, perl=TRUE)
  clean_text <- gsub(pattern=edcs_number_three_right[1], replacement=edcs_number_three_right[2], x=clean_text, perl=TRUE)
  clean_text <- gsub(pattern=edcs_number_three_left[1], replacement=edcs_number_three_left[2], x=clean_text, perl=TRUE)
  clean_text <- gsub(pattern=edcs_number_six_both[1], replacement=edcs_number_six_both[2], x=clean_text, perl=TRUE)
  clean_text <- gsub(pattern=edcs_number_one[1], replacement=edcs_number_one[2], x=clean_text, perl=TRUE)
  clean_text <- gsub(pattern=edcs_quotes[1], replacement=edcs_quotes[2], x=clean_text, perl=TRUE)
  clean_text <- gsub(pattern=edcs_backslashes[1], replacement=edcs_backslashes[2], x=clean_text, perl=TRUE)
  clean_text <- gsub(pattern=expanded_abbreviations_conservative[1], replacement=expanded_abbreviations_conservative[2], x=clean_text, perl=TRUE)
  clean_text <- gsub(pattern=suppresion_keep_interpretive[1], replacement=suppresion_keep_interpretive[2], x=clean_text, perl=TRUE)
  clean_text <- gsub(pattern=restoration_conservative[1], replacement=restoration_conservative[2], x=clean_text, perl=TRUE)
  clean_text <- gsub(pattern=substitution_edh_conservative[1], replacement=substitution_edh_conservative[2], x=clean_text, perl=TRUE)
  clean_text <- gsub(pattern=substitution_edh_conservative_missing[1], replacement=substitution_edh_conservative_missing[2], x=clean_text, perl=TRUE)
  clean_text <- gsub(pattern=substitution_conservative[1],replacement=substitution_conservative[2], x=clean_text, perl=TRUE)
  clean_text <- gsub(pattern=new_line[1], replacement=new_line[2], x=clean_text, perl=TRUE)
  clean_text <- gsub(pattern=interpunction_symbols[1], replacement=interpunction_symbols[2], x=clean_text, perl=TRUE)
  clean_text <- gsub(pattern=epigraphic_symbols[1], replacement=epigraphic_symbols[2], x=clean_text, perl=TRUE)
  clean_text <- gsub(pattern=uncertainty_symbols[1], replacement=uncertainty_symbols[2], x=clean_text, perl=TRUE)
  clean_text <- gsub(pattern=arabic_numerals[1], replacement=arabic_numerals[2], x=clean_text, perl=TRUE)
  clean_text <- gsub(pattern=unclosed_brackets[1], replacement=unclosed_brackets[2], x=clean_text, perl=TRUE)
  clean_text <- gsub(pattern=edcs_que[1], replacement=edcs_que[2], x=clean_text, perl=TRUE)
  clean_text <- gsub(pattern=edcs_vir[1], replacement=edcs_vir[2], x=clean_text, perl=TRUE)
  clean_text <- gsub(pattern=extra_blank[1], replacement=extra_blank[2], x=clean_text, perl=TRUE)
  clean_text <- gsub(pattern=multi_whitespace[1], replacement=multi_whitespace[2], x=clean_text, perl=TRUE)
  clean_text <- gsub(pattern=whitespace_endline[1], replacement=whitespace_endline[2], x=clean_text, perl=TRUE)
  return(clean_text)
}
```




### Interpretive model for text mining analysis based on the analysis of words
*Aim:* to have a clean text enriched by editorial interpretations and reconstructions of the text (to have as rich text of an inscription as possible).

The output of the function will consist of words separated by one space, so the data is ready for tidytext analysis. No interpunction will be left in the text.

```{r}
cleaning_interpretive_edcs <- function(epigraphic_dataset){
  clean_text <- gsub(pattern=dubious_dot_subscript[1], replacement=dubious_dot_subscript[2], x=epigraphic_dataset, perl=TRUE)
  clean_text <- gsub(pattern=edcs_number_three_both[1], replacement=edcs_number_three_both[2], x=clean_text, perl=TRUE)
  clean_text <- gsub(pattern=edcs_number_three_right[1], replacement=edcs_number_three_right[2], x=clean_text, perl=TRUE)
  clean_text <- gsub(pattern=edcs_number_three_left[1], replacement=edcs_number_three_left[2], x=clean_text, perl=TRUE)
  clean_text <- gsub(pattern=edcs_number_six_both[1], replacement=edcs_number_six_both[2], x=clean_text, perl=TRUE)
  clean_text <- gsub(pattern=edcs_number_one[1], replacement=edcs_number_one[2], x=clean_text, perl=TRUE)
  clean_text <- gsub(pattern=edcs_quotes[1], replacement=edcs_quotes[2], x=clean_text, perl=TRUE)
  clean_text <- gsub(pattern=edcs_backslashes[1], replacement=edcs_backslashes[2], x=clean_text, perl=TRUE)
  clean_text <- gsub(pattern=expanded_abbreviations_interpretive[1], replacement=expanded_abbreviations_interpretive[2], x=clean_text, perl=TRUE)
  clean_text <- gsub(pattern=suppresion_remove_interpretive[1], replacement=suppresion_remove_interpretive[2], x=clean_text, perl=TRUE)
  clean_text <- gsub(pattern=restoration_interpretive[1], replacement=restoration_interpretive[2], x=clean_text, perl=TRUE)
  clean_text <- gsub(pattern=substitution_edh_interpretive[1], replacement=substitution_edh_interpretive[2], x=clean_text, perl=TRUE)
  clean_text <- gsub(pattern=substitution_edh_interpretive_missing[1], replacement=substitution_edh_interpretive_missing[2], x=clean_text, perl=TRUE)
  clean_text <- gsub(pattern=substitution_interpretive[1],replacement=substitution_interpretive[2], x=clean_text, perl=TRUE)
  clean_text <- gsub(pattern=new_line[1], replacement=new_line[2], x=clean_text, perl=TRUE)
  clean_text <- gsub(pattern=interpunction_symbols[1], replacement=interpunction_symbols[2], x=clean_text, perl=TRUE)
  clean_text <- gsub(pattern=epigraphic_symbols[1], replacement=epigraphic_symbols[2], x=clean_text, perl=TRUE)
  clean_text <- gsub(pattern=uncertainty_symbols[1], replacement=uncertainty_symbols[2], x=clean_text, perl=TRUE)
  clean_text <- gsub(pattern=arabic_numerals[1], replacement=arabic_numerals[2], x=clean_text, perl=TRUE)
  clean_text <- gsub(pattern=unclosed_brackets[1], replacement=unclosed_brackets[2], x=clean_text, perl=TRUE)
  clean_text <- gsub(pattern=edcs_que[1], replacement=edcs_que[2], x=clean_text, perl=TRUE)
  clean_text <- gsub(pattern=edcs_vir[1], replacement=edcs_vir[2], x=clean_text, perl=TRUE)
  clean_text <- gsub(pattern=extra_blank[1], replacement=extra_blank[2], x=clean_text, perl=TRUE)
  clean_text <- gsub(pattern=multi_whitespace[1], replacement=multi_whitespace[2], x=clean_text, perl=TRUE)
  clean_text <- gsub(pattern=whitespace_endline[1], replacement=whitespace_endline[2], x=clean_text, perl=TRUE)
  return(clean_text)
}
```


### Enriching the full dataset with conservative and interpretive cleaned versions of the text:
```{r}
EDCS_clean <- EDCS_clean %>%
   mutate(clean_text_conservative = cleaning_conservative_edcs(EDCS_clean$inscription_stripped_final)) %>% 
   mutate(clean_text_interpretive_word = cleaning_interpretive_edcs(EDCS_clean$inscription_stripped_final))
  
```


### Combining notes extracted from the text of the inscription by their category:
```{r}
EDCS_clean <- EDCS_clean %>%
  mutate(notes_dating = paste(EDCS_clean$dating_notes, EDCS_clean$dating_notes2))%>%
  mutate(notes_dating = str_replace_all(notes_dating, "NA", "")) %>%
  mutate(notes_references = paste(EDCS_clean$other_notes, EDCS_clean$other_notes2))%>%
  mutate(notes_references = str_replace_all(notes_references, "NA", "")) %>%
  mutate(notes_comments = paste(EDCS_clean$other_notes3, EDCS_clean$other_notes4, EDCS_clean$other_notes5, EDCS_clean$other_notes6, EDCS_clean$other_notes7, EDCS_clean$other_notes8)) %>%
  mutate(notes_comments = str_replace_all(notes_comments, "NA", ""))
```

```{r}
names(EDCS_clean)
```


### Selecting the columns relevant for further analysis (for export)
```{r}
EDCS_export <- EDCS_clean %>%
  select(`EDCS-ID`, publication, province, province_list , place, place_list, Links, language, `dating from`, `dating to`, start_yr, end_yr_list, end_yr_1, notes_dating, status, status_list, inscr_type, status_notation, inscr_process, Latitude, Longitude, photo, Material, Comment, notes_references, notes_comments, inscription, inscription_stripped_final, clean_text_interpretive_word, clean_text_conservative, notes_dating, notes_references, notes_comments)
```



# Saving to Sciencedata
```{r}
EDCS_cleaned_json <- jsonlite::toJSON(EDCS_export, auto_unbox = TRUE)

#mycred_secret<- readLines("~/mysecret.txt")
write(EDCS_cleaned_json, file="../output/EDCS_text_cleaned_2022-02-15.json")
#request("output/EDCS_text_cleaned_2021-03-01.json", path="/sharingout/648597@au.dk/SDAM_root/SDAM_data/EDCS/public",
        #method="PUT", cred=c(mycred_secret[1], mycred_secret[2])) # does not work any more, has to be uploaded manually to sciencedata.dk
```

# Remove local copy of the json and the credential before submitting to GitHub
```{r}
#file.remove("./EDCS_text_cleaned_2020-10-09.json")
#remove(mycred_secret)
```
