---
title: "EDCS cleaning of the epigraphic text"
author: "Petra Hermankova"
date: "10/12/2020"
output:
  html_document:
    theme: united
    toc: yes
    toc_float: true
    number_sections: true
    toc_depth: 2
    df_print: paged
---

```{r setup, include=FALSE, echo=FALSE}
require("knitr")

library(tidyverse)
library(jsonlite)
library(leaflet)
```

## Loading data
Load the dataset, if you have Sciencedata.dk credentials

```{r}
EDCS <- jsonlite::fromJSON("https://sciencedata.dk/public/1f5f56d09903fe259c0906add8b3a55e/EDCS_text_cleaned_2022-02-15.json")
```

```{r}
EDCS <- jsonlite::fromJSON("../output/EDCS_text_cleaned_2022-02-15.json")
```


# Dataset exploration
```{r}
names(EDCS)
```

```{r}
head(EDCS)
```

## Some of the records are empty

```{r}
EDCS_empty<- EDCS %>% 
  filter(is.na(`EDCS-ID`))
```

## Non-empty records
```{r}
EDCS<- EDCS %>% 
  filter(!is.na(`EDCS-ID`))
```

## How many non empty inscriptions there are in total?
```{r}
nrow(EDCS)
```


## Language

### How many and which combinations of languages contain inscription in Greek?
```{r}
grep(x = EDCS$language, "GR", value = TRUE) %>%
  table()
```

### How many inscriptions have GR as one of their languages?
```{r}
grep(x = EDCS$language, "GR", value = TRUE) %>%
  length() -> greek

greek
greek/(nrow(EDCS)/100)
```

### How many and which combinations of languages are in EDCS?
```{r}
unique(EDCS$language)
```
```{r}
table(EDCS$language)
```
### How much inscriptions are in Latin
```{r}
latin<- sum(is.na(EDCS$language))
latin
latin/(nrow(EDCS)/100)
```


## Province
```{r}
EDCS %>%
  count(province, sort = TRUE)
```

Some inscriptions have double provinces, e.g.:

Belgica | Germania inferior
Pannonia inferior | Pannonia superior
Moesia inferior | Moesia superior

```{r}
str_split_fixed(EDCS$province, pattern = " \\| ", n=2) %>% as.data.frame() -> provinces

provinces %>%
  cbind(combined = c(provinces$V1, provinces$V2)) %>%
  filter(combined != "") -> provinces

provinces %>%
  count(combined, sort=TRUE)
```
```{r}
location %>% 
  filter(province == "Sicilia" & !is.na(start_yr) & !is.na(Latitude)) -> sicilia
```

```{r}
sicilia %>% 
  filter(!is.na(clean_text_interpretive_word))
```


```{r, eval=FALSE}
map_sicilia <- leaflet(width="100%") %>%
 addProviderTiles("Stamen.TerrainBackground")%>% # Add CartoDB map tiles
# addProviderTiles("Stamen.Watercolor")%>% # Add CartoDB map tiles
# addProviderTiles("Esri.WorldImagery", group = "ESRI Aerial") %>%
# addPolylines(data = roads, color = "purple", weight = 2, opacity = 0.7) %>%
  setView( lng = 22.326743, lat = 46.897122, zoom = 4 ) %>%
  #setMaxBounds(lat1=43.633977, lng1 =-11.227926 , lat2=35.133882 , lng2=50.882336) %>%
  addCircles(lng = sicilia$Longitude, lat = sicilia$Latitude, opacity = 0.5, radius = 10, fill = TRUE, color = "grey" , fillColor = "white",) %>%
  addCircles(lng = sicilia$Longitude,  
             lat = sicilia$Latitude, opacity = 0.5, radius = 10, fill = TRUE, color = "purple" , fillColor = "",
             popup = paste0("<b> InscriptionID: </b>", sicilia$ID
                             # "<br><b> Province: </b>", sicilia$province,
                             #<br><b> Type of inscription: </b>", lat_long_milestone$type_of_inscription_clean,
                             # "<br><b> Type of monument: </b>", lat_long_milestone$type_of_monument_clean,
                             # "<br><b> Not before (date): </b>", lat_long_milestone$not_before,
                             # "<br><b> Not after (date): </b>", lat_long_milestone$not_after,
                             #   "<br><b> Commentary: </b>", lat_long_milestone$commentary,
                            )
             ) %>%
addLegend(position = "topright",
  colors = c("Purple"),
  labels = c("Inscription"), opacity = 1,
  title = "Inscriptions in the EDCS database with known coordinates"
)
map_sicilia
```


## Place
```{r}
# unique place names
length(unique(unlist(EDCS$place_list)))
```

```{r}
#first element of the place list
plc_1<- map(EDCS$place_list,1)

plc1_df <- plc_1 %>% 
  unlist() 

plc1_df <- as.data.frame(plc1_df)
plc1_df %>% 
  count(plc1_df, sort=T)
```

```{r}
place<- unnest_longer(EDCS, col=place_list, simplify = T)
```

```{r}
place %>% 
  count(place_list, sort=T)
```



# How many inscriptions are linked to EDH

```{r}
heidelberg_data<- length(str_subset(EDCS$Links, "heidelberg"))
heidelberg_data

heidelberg_data/(nrow(EDCS)/100)
```

# How many inscriptions are linked to PHI

```{r}
length(str_subset(EDCS$Links, "P\\d{1,6}"))
```

# Status attribute

### Type of inscription
```{r}
inscr_type<- unlist(EDCS$inscr_type) 
  
inscr_type_df <- as.data.frame(inscr_type) %>% count(inscr_type, sort = T)
inscr_type_df
```



### Type of inscribing process
```{r}
inscr_process<- unlist(EDCS$inscr_process) 
inscr_process_df <- as.data.frame(inscr_process) %>% count(inscr_process, sort = T)
inscr_process_df
```


### Type of status notation
```{r}
status_notation<- unlist(EDCS$status_notation) 
status_notation_df <- as.data.frame(status_notation) %>% count(status_notation, sort = T)
status_notation_df
```

## Coordinates

### Which inscriptions have geospatial coordinates?
```{r}
EDCS %>%
  dplyr::filter(!is.na(Latitude)) -> location
# how many inscriptions have latitude
nrow(location)
# how many % of inscriptions have latitude
nrow(location)/(nrow(EDCS)/100)
```

```{r}
location$Latitude <- as.numeric(location$Latitude)
location$Longitude <- as.numeric(location$Longitude)
```

```{r}
location %>% 
  rename(EDCS_ID = `EDCS-ID`)
```

### How precise are the coordinates (how many decimal numbers have) 
```{r}
decimalplaces<-function(x) {
  x_str <- toString(x)
  x_str<- str_split(x_str, "\\.")
  nchar(x_str[[1]][2])
}

### latitude
lat_df<- as.data.frame(EDCS$Latitude) %>%  
  mutate(decimals = unlist(map(EDCS$Latitude, decimalplaces)))

lat_df_count<- lat_df %>% 
  count(decimals, sort = T) %>% 
  mutate(percent = n/(sum(n)/100))

lat_df_count

### longitude
long_df<- as.data.frame(EDCS$Longitude) %>%  
  mutate(decimals = unlist(map(EDCS$Longitude, decimalplaces)))

long_df_count<- long_df %>% 
  count(decimals, sort = T) %>% 
  mutate(percent = n/(sum(n)/100))

long_df_count
```


### Mapped inscriptions with known coordinates - for full dataset be careful before running
```{r, eval=FALSE}
map_all <- leaflet(width="100%") %>%
 addProviderTiles("Stamen.TerrainBackground")%>% # Add CartoDB map tiles
# addProviderTiles("Stamen.Watercolor")%>% # Add CartoDB map tiles
# addProviderTiles("Esri.WorldImagery", group = "ESRI Aerial") %>%
# addPolylines(data = roads, color = "purple", weight = 2, opacity = 0.7) %>%
  setView( lng = 22.326743, lat = 46.897122, zoom = 4 ) %>%
  #setMaxBounds(lat1=43.633977, lng1 =-11.227926 , lat2=35.133882 , lng2=50.882336) %>%
#  addCircles(lng = location$Longitude, lat = location$Latitude, opacity = 0.5, radius = 10, fill = TRUE, color = "grey" , fillColor = "white",) %>%
  addCircles(lng = location$Longitude,  
             lat = location$Latitude, opacity = 0.5, radius = 10, fill = TRUE, color = "purple" , fillColor = "",
             popup = paste0("<b> InscriptionID: </b>", location$ID
                             # "<br><b> Province: </b>", location$province,
                             #<br><b> Type of inscription: </b>", lat_long_milestone$type_of_inscription_clean,
                             # "<br><b> Type of monument: </b>", lat_long_milestone$type_of_monument_clean,
                             # "<br><b> Not before (date): </b>", lat_long_milestone$not_before,
                             # "<br><b> Not after (date): </b>", lat_long_milestone$not_after,
                             #   "<br><b> Commentary: </b>", lat_long_milestone$commentary,
                            )
             ) %>%
addLegend(position = "topright",
  colors = c("Purple"),
  labels = c("Inscription (n = 486,699)"), opacity = 1,
  title = "Inscriptions in the EDCS database with known coordinates"
)
# map_all
```



# Dates

## Start year
```{r}
EDCS %>%
  filter(!is.na(EDCS$start_yr)) -> has_start_yr
# how many inscriptions have start year
nrow(has_start_yr)
# how many inscriptions have start year in %
nrow(has_start_yr)/(nrow(EDCS)/100)
```

## End year
```{r}
EDCS %>%
  filter(!is.na(EDCS$end_yr_1)) -> has_end_yr
# how many inscriptions have end year
nrow(has_end_yr)
# how many inscriptions have end year in %
nrow(has_end_yr)/(nrow(EDCS)/100)
```

### Date Range
```{r}
has_end_yr<- has_end_yr %>% 
  mutate(date_interval = (start_yr - end_yr_1) * -1)

table(has_end_yr$date_interval)
summary(has_end_yr$date_interval)

date_int<- has_end_yr %>% 
  count(date_interval, sort=T) %>% 
  #filter(date_interval > 0 & date_interval < 1000) %>% 
  mutate(freq = n/(nrow(has_end_yr)/100))
  
date_int %>% 
  ggplot(aes(x=date_interval,)) + 
  geom_histogram(fill="blue", color="blue", alpha=0.5, size=0.2) +
  labs(x = "Length of dating interval", y = "%", title ="Length of dating intervals in years in EDCS dataset") +
  geom_vline(aes(xintercept=100))+
  coord_cartesian(xlim = c(0, 1000), ylim=c(0,100))

```

```{r}
table(date_int)
```



```{r}
summary(EDCS$start_yr)

plot(table(EDCS$start_yr))
```

```{r}
summary(EDCS$end_yr_1)
plot(table(EDCS$end_yr_1))
```

### Both dates
```{r}
EDCS %>% 
  filter(!is.na(start_yr) & !is.na(end_yr_1)) -> has_both_dates

has_both_dates
nrow(has_both_dates)/(nrow(EDCS)/100)
```


# Material

```{r}
EDCS %>%
  count(Material, sort= TRUE)
```

# Comments

First 20 Comments to see the nature of the contents:
```{r}
unique(EDCS$Comment)[1:20]
```




## EDCS without EDH

```{r}
EDCScon <- jsonlite::fromJSON("../data/EDCScon_dated.json")
EDCScon<- as_tibble(EDCScon)
```

```{r}
EDCSconjson <- jsonlite::toJSON(EDCScon, auto_unbox = TRUE)
EDCScon_json<- jsonlite::fromJSON(EDCSconjson)
EDCScon<- as_tibble(EDCScon_json)
head(EDCScon,2)
```

```{r}
library(stringr)

length(str_subset(EDCS$Links, "trismegistos"))
```

```{r}
length(str_subset(EDCS$Links, "edr"))
```
```{r}
length(str_subset(EDCS$Links, "edh"))
```


  
```{r}
table(EDCS$Links)
```
  

```{r}
EDCS %>% 
  dplyr::filter(is.na(EDCS$Links) & !is.na(EDCS$Latitude))

EDCS %>% 
  dplyr::filter(is.na(EDCS$Latitude))
```


```{r}
filter(EDCS$Links = "list()")
```

